Improved Bayesian Ethnorace Prediction

### 2021-02-11 paper.Rmd
more!

```{r, fig.cap="Accuracy Scores by Input Data Availability"}
acc_bper <- tests %>% 
  filter(test_method == "bper") %>% 
  select(test_type, accuracy)

ggplot(acc_bper, aes(x = test_type, y = accuracy, group = 1)) +
  geom_line(color = "darkcyan") +
  geom_point(size = 3, color = "darkcyan") +
  theme_minimal() +
  ylim(0.75, 0.85) +
  labs(y = "Accuracy Score") +
  scale_x_discrete(name = "", limits = c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name",
                     "zip" = "ZIP \n Party ID \n First Name \n Last Name",
                     "zip_np" = "ZIP \n First Name \n Last Name",
                     "state" = "State \n Party ID \n First Name \n Last Name",
                     "state_np" = "State \n First Name \n Last Name"
                   )) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7))
```

As expected, the model with the greatest number of input data sources, and at the most precise geographic level, on the far left of the figure performs the best in terms of overall accuracy. Using Census blocks, multi-unit address, party ID, first names, and last names, this model correctly identifies the ethnorace for `r subset(tests, select = "accuracy", test_method == "bper" & test_type == "all") * 100`% of individuals in the sample. The downward trend in model accuracy seen in the figure corresponds in expanding the size of the geolocation variable. Moving from blocks to ZIP codes, to places, to counties, and to states all decrease the overall predictive performance. I include the input for multi-unit occupancy only for the Census block models because I believe this reflects the practical contexts where `bper` might be used. If a researcher has access to individual-level addresses, they should be able to using geocoding to find the matching Census blocks and should be able to parse the unit type. But researchers relying on more aggregate geographies likely do not have access to individual addresses, and hence the unit types, for their sample.[^11] In Figure 2 I also pair each geolocation variable with a model missing the party ID input. The blow to predictive performance across all geographies appears roughly uniform with the removal of party ID.

@@

```{r, fig.cap="Precision/Recall Scores by Ethnorace and Input Data: `bper`"}
metrics_bper <- tests %>% 
  filter(test_method == "bper") %>% 
  select(test_type, contains(c("prec", "rec"))) %>% 
  pivot_longer(cols = contains(c("prec", "rec"))) %>% 
  separate(name, c("Metric", "Race"), sep = "_") %>% 
  filter(Race != "other") %>% 
  mutate(Metric = case_when(Metric == "prec" ~ "Precision",
                            Metric == "rec" ~ "Recall"),
         Race = case_when(Race == "api" ~ "Asian",
                          Race == "black" ~ "Black",
                          Race == "hispanic" ~ "Hispanic",
                          Race == "white" ~ "White"))

ggplot(metrics_bper, aes(x = test_type, y = value, group = Metric)) +
  geom_line(aes(color = Metric)) +
  geom_point(aes(color = Metric), size = 1.5) +
  theme_minimal() +
  labs(x = "", y = "Score") +
  scale_color_manual(values = c("gold3", "skyblue4")) +
  scale_x_discrete(name = "", limits = c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np"),
                   labels = c(
                     "all" = "Block", 
                     "place" = "Place", 
                     "county" = "County ", 
                     "all_np" = "Block (np)", 
                     "place_np" = "Place (np)", 
                     "county_np" = "County (np)",
                     "zip" = "ZIP",
                     "zip_np" = "ZIP (np)",
                     "state" = "State",
                     "state_np" = "State (np)"
                   )) +
  facet_wrap(~ Race) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7, angle = 60))
```

@@

ggplot(accuracy, aes(x = test_type, y = diff)) +
  geom_bar(stat = "identity", fill = "darkcyan") +
  theme_minimal() + 
  labs(y = "Difference in Accuracy", x = "") +
  scale_x_discrete(name = "", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))
```

@@

rec_api <-
 ggplot(api, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))

@@

rec_black <-
 ggplot(black, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(-0.3, 0) +
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))

@@

rec_hispanic <-
 ggplot(hispanic, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))

@@

rec_white <-
 ggplot(white, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))

>>>>

```{r, fig.cap="Accuracy Scores by Input Data"}
acc_bper <- tests %>% 
  filter(test_method == "bper") %>% 
  select(test_type, accuracy)

ggplot(acc_bper, aes(x = test_type, y = accuracy, group = 1)) +
  geom_line(color = "darkcyan") +
  geom_point(size = 3, color = "darkcyan") +
  theme_minimal() +
  ylim(0.75, 0.85) +
  labs(y = "Accuracy Score") +
  scale_x_discrete(name = "Model Inputs", limits = c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name",
                     "zip" = "ZIP \n Party ID \n First Name \n Last Name",
                     "zip_np" = "ZIP \n First Name \n Last Name",
                     "state" = "State \n Party ID \n First Name \n Last Name",
                     "state_np" = "State \n First Name \n Last Name"
                   )) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7))
```

As expected, the model with the greatest number of input data sources, and at the most precise geographic level, on the far left of the figure performs the best in terms of overall accuracy. Using Census blocks, multi-unit address, party ID, first names, and last names, this model correctly identifies the ethnorace of `r round(subset(tests, select = "accuracy", test_method == "bper" & test_type == "all"), 3) * 100`% of individuals in the sample. The downward trend in model accuracy seen in the figure corresponds to expanding the size of the geolocation variable. Moving from blocks to ZIP codes, to places, to counties, and to states all decrease the overall predictive performance. I include the input for multi-unit occupancy only for the Census block models because I believe this reflects the practical contexts where `bper` might be used. If a researcher has access to individual-level addresses, they should be able to using geocoding to find the matching Census blocks and should be able to parse the unit type. But researchers relying on more aggregate geographies likely do not have access to individual addresses, and hence the unit types, for their sample.[^11] In Figure 2 I also pair each geolocation variable with a model missing the party ID input. The blow to predictive performance across all geographies appears roughly uniform with the removal of party ID.

@@

Accuracy scores, however, are another incomplete metric for assessing predictive performance. In contexts where the true distribution of classes is highly imbalanced, Accuracy can provide overly-optimistic results. For example, if we were to simply classify every individual as White in the North Carolina/Florida voter file, we would achieve 65% Accuracy without even trying! We can evaluate the models in a more rigorous way by looking at each ethnorace category separately.

```{r, fig.cap="Precision/Recall Scores by Ethnorace and Input Data"}
metrics_bper <- tests %>% 
  filter(test_method == "bper") %>% 
  select(test_type, contains(c("prec", "rec"))) %>% 
  pivot_longer(cols = contains(c("prec", "rec"))) %>% 
  separate(name, c("Metric", "Race"), sep = "_") %>% 
  filter(Race != "other") %>% 
  mutate(Metric = case_when(Metric == "prec" ~ "Precision",
                            Metric == "rec" ~ "Recall"),
         Race = case_when(Race == "api" ~ "Asian",
                          Race == "black" ~ "Black",
                          Race == "hispanic" ~ "Hispanic",
                          Race == "white" ~ "White"))

ggplot(metrics_bper, aes(x = test_type, y = value, group = Metric)) +
  geom_line(aes(color = Metric)) +
  geom_point(aes(color = Metric), size = 1.5) +
  theme_minimal() +
  labs(x = "", y = "Score") +
  scale_color_manual(values = c("gold3", "skyblue4")) +
  scale_x_discrete(name = "Model Inputs", limits = c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np"),
                   labels = c(
                     "all" = "Block", 
                     "place" = "Place", 
                     "county" = "County ", 
                     "all_np" = "Block (np)", 
                     "place_np" = "Place (np)", 
                     "county_np" = "County (np)",
                     "zip" = "ZIP",
                     "zip_np" = "ZIP (np)",
                     "state" = "State",
                     "state_np" = "State (np)"
                   )) +
  facet_wrap(~ Race) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7, angle = 60))
```
Figure 3 displays the Precision and Recall scores broken down by race for the same models used in Figure 2. Precision is the percentage of individuals who  of correctly classified individuals in a single class.

@@

ggplot(accuracy, aes(x = test_type, y = diff)) +
  geom_bar(stat = "identity", fill = "darkcyan") +
  theme_minimal() + 
  labs(y = "Difference in Accuracy", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))
```

@@

rec_api <-
 ggplot(api, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))

@@

rec_black <-
 ggplot(black, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  ylim(-0.3, 0) +
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))

@@

rec_hispanic <-
 ggplot(hispanic, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))

@@

rec_white <-
 ggplot(white, aes(x = test_type, y = rec_diff, fill = rec_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity(guide = F) +
  theme_minimal() + 
  labs(y = "Difference in Recall", x = "") +
  scale_x_discrete(name = "Model Inputs", limits = c("all", "all_np", "place", "place_np", "county", "county_np"),
                   labels = c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name")) +
  theme(text = element_text(family = "serif"))
